Article figures

### 2020-11-02 src/article_figures.Rmd
Improved article figures

---
title: "Article figures"
author: "Jarkko Toivonen"
date: "10/30/2020"
output: html_document
---

@@

@@

```{r}
# Creates a forest plot of means and confidence intervals from posterior distributions.
# Puts both male and female results in the same plot.
create_double_forest_plot <- function(male_posterior, female_posterior, variables, combined_variables, base_size = 11) {
  for (gender in c("male", "female")) {
    #posterior <- ifelse(gender == "male", male_posterior, female_posterior)
    if (gender == "male") {
      posterior <- male_posterior
    } else {
      posterior <- female_posterior
    }

    for (i in seq_along(posterior)) {
      #cat(sprintf("Gender %s, Column %i\n", gender, i))
      v <- posterior[[i]]
      ci_hdi <- bayestestR::ci(v, method = "HDI", ci=0.95)
      #str(ci_hdi)
      L <- list(names = variables[[i]], gender=gender, mean = mean(v), low = ci_hdi$CI_low, high = ci_hdi$CI_high)
      if (i == 1 && gender=="male") {
        result <- data.frame(L, stringsAsFactors = FALSE)
      } else {
        result <- rbind(result, L)
      }
    }
  }
  result <- as_tibble(result)
  result <- result %>%
    mutate(gender = factor(gender, levels=c("female", "both", "male")))
  cis <- result
  #result <- DescTools::Rev(result, margin=1)   # Reverse the order of rows so they appear in correct order in the forest plot
  print(head(result))
  combined_variables <- combined_variables %>% rename(names=Pretty)
  result <- left_join(combined_variables, result, by="names")
  plot <- result %>% ggplot(aes(y=factor(names, levels=rev(combined_variables$names)), x=mean, xmin=low, xmax=high, color=gender)) +     
    geom_vline(aes(xintercept=0), color="lightgray") +
    #geom_pointrange() +
    ggstance::geom_pointrangeh(position=position_dodge2(width=0.8, padding=0.1), size=0.25) + # ggstance is required to make legend keys horizontal
    labs(title="", #title="Effects sizes of variables on Hb prediction",
         x="Regression coefficient", y=NULL, colour="Sex") +
    scale_colour_discrete(drop=FALSE, breaks = c("male", "female")) + 
    #guides(colour = guide_legend(reverse = TRUE)) + # Reverse the order of keys in the legend
    theme_gray(base_size = base_size) + theme(legend.position = "bottom", legend.direction = "horizontal")

  return(list(plot=plot, cis=cis))
}

@@

@@

  #pretty <- descript["Pretty"]
  pretty <- combined_variables
  rrfFit.varimp <- left_join(pretty, rrfFit.varimp, by="Pretty")
  varimp.plot <- rrfFit.varimp %>% 
    ggplot(aes(y=factor(Pretty, levels=rev(Pretty)), x=Importance, xmin=0, xmax=Importance, colour = factor("both", levels=c("female", "both", "male")))) + 
    ggstance::geom_pointrangeh(position=position_dodge2(width=0.8, padding=0.1)) +  # ggstance is required to make legend keys horizontal
                               #colour = hue_pal()(3)[2], size=0.25) +
    labs(y=NULL, colour="Sex") + 
    guides(colour="none", y="none") +
    scale_colour_discrete(drop=FALSE) +
    theme_gray(base_size = base_size)

@@

create_forest_importance_plot <- function(male_posterior, female_posterior, variables,
                                           filename=NULL,
                                           width = 180,  # width of the combined figure in millimetres
                                          base_size = 11
) {
  rrfFit_roc <- load_single("~/FRCBS/interval_prediction/results/rrfFit_roc.rdata")
  combined_variables <- tibble(Pretty=c(variables, "Life time donations", "Sex", "First Hb"))
  forest <- create_double_forest_plot(male_posterior, female_posterior, variables, combined_variables, base_size = base_size)$plot
  importance <- create_variable_importance_plot2(rrfFit_roc, descript, combined_variables, base_size = base_size)

@@

```{r}
finngenn_male_raw <- get_raw_result_list("finngen-male-dlmm")
finngenn_female_raw <- get_raw_result_list("finngen-female-dlmm")
male_posterior <- finngenn_male_raw$samples
female_posterior <- finngenn_female_raw$samples
variables <- finngenn_male_raw$pretty_variable_names
```

@@

  g <- df %>% ggplot(aes(y=Id, x=value, xmin=low, xmax=high)) + 
    #ggstance::geom_pointrangeh(aes(colour=sex, shape=model)) + 
    geom_pointrange(aes(colour=sex, fill=sex, shape=model), key_glyph = draw_key_rect) + 
    labs(x="Value", y="Data and method", colour="Sex", shape="Model") + 
    scale_colour_discrete(breaks = c("male", "female", "both")) +
    scale_shape_discrete(breaks=c("lmm", "dlmm", "dt", "rf")) +
    #scale_colour_manual(values = c("female" = 1, "male" = 2, "both" = 3)) +
    geom_blank(data=dummy) +
    facet_wrap("type", scales="free_x")
  #ggsave(filename="~/FRCBS/results-for-eba-paper/forest_plot.pdf", plot=g, dpi=600, units="mm", width=180)
  return(g)
}

@@

>>>>

---
title: "Article figures"
author: "Jarkko Toivonen"
date: "10/30/2020"
output: 
  html_document: default
  md_document: default
---

@@

descript <- tibble(Variable = c("donor", "Hb", "days_to_previous_fb", "age", "previous_Hb_def", 
                                "year", "warm_season", "consecutive_deferrals", "recent_donations",
                                "recent_deferrals", "hour", 
                                "previous_Hb", "Hb_first", "Hb_deferral","nb_donat","gender"), 
                   Pretty = c("Donor ID", "Hemoglobin", "Days to previous full blood donation", "Age", 
                              "Previous donation deferred", #"Previous Hb deferral",   # This is a more descriptive name
                              "Year", "Warm season", "Consecutive deferrals", "Recent donations", 
                              "Recent deferrals", "Hour", 
                              "Previous Hb", "First Hb", "Hb deferral","Life time donations","Sex"),
                   Type = c("Factor", "numeric", "numeric (int)", "numeric", "boolean",
                            "numeric (int)", "boolean", "numeric (int)", "numeric (int)", "numeric (int)", "numeric",
                            "numeric", "numeric", "boolean","numeric (int)","Factor"),
                   Explanation = c("Donor identifier",
                                   "Amount of Hemoglobin",
                                   "Time (in days) between Hb measurement and previous full blood donation event",
                                   "Age of donor",
                                   "Indicates whether the donor was deferred from blood donation due to low hemoglobin at previous donation event",
                                   "Year of donation",
                                   "True if donation was given in April-September",
                                   "Amount of times the donor has been deferred due to low hemoglobin since last succesful whole blood donation",
                                   "Amount of donations in the last two years",
                                   "Amount of deferrals due to low hemoglobin in the last two years",
                                   "Time of day when donation was given as hours (e.g. 13:45 = 13.75)",
                                   "Hb value at previous measurement (ICP-model)",
                                   "Hb value at first donation of this donor (Non ICP-model)",
                                   "Deferred based on low hemoglogin",
                                  "Life time total count of donations",
                                  "Sex"
                   )
                   )

@@

@@

# Creates a forest plot of means and confidence intervals from posterior distributions.
# Puts both male and female results in the same plot.
create_double_forest_plot <- function(male_posterior, female_posterior, variables, combined_variables, base_size = 11) {
  for (gender in c("male", "female")) {
    #posterior <- ifelse(gender == "male", male_posterior, female_posterior)
    if (gender == "male") {
      posterior <- male_posterior
    } else {
      posterior <- female_posterior
    }

    for (i in seq_along(posterior)) {
      #cat(sprintf("Gender %s, Column %i\n", gender, i))
      v <- posterior[[i]]
      ci_hdi <- bayestestR::ci(v, method = "HDI", ci=0.95)
      #str(ci_hdi)
      L <- list(names = variables[[i]], gender=gender, mean = mean(v), low = ci_hdi$CI_low, high = ci_hdi$CI_high)
      if (i == 1 && gender=="male") {
        result <- data.frame(L, stringsAsFactors = FALSE)
      } else {
        result <- rbind(result, L)
      }
    }
  }
  result <- as_tibble(result)
  result <- result %>%
    mutate(gender = factor(gender, levels=c("female", "both", "male"))) %>%
    rename(Pretty=names) %>%
    mutate(Pretty=factor(Pretty, levels=levels(combined_variables$Pretty)))
  cis <- result
  #result <- DescTools::Rev(result, margin=1)   # Reverse the order of rows so they appear in correct order in the forest plot
  print(head(result))
  #combined_variables <- combined_variables %>% 
    #rename(names=Pretty) %>%
    #mutate(stripe=factor(row_number()%%2))
  result <- left_join(combined_variables, result, by="Pretty") %>%
    mutate(Pretty=fct_rev(Pretty))

    #mutate(Pretty=factor(Pretty, levels=rev(combined_variables$Pretty)))
    # group_by(gender) %>%
    # mutate(stripe=factor(row_number()%%2)) %>%
    # ungroup()
  df1 <<- result
  plot <- result %>% 
    # mutate(Pretty=factor(names, levels=rev(combined_variables$names)), 
    #        stripe=factor(row_number()%%2)) %>%
    ggplot() +     
    geom_vline(aes(xintercept=0), color="lightgray") +
    #geom_pointrange() +
    ggstance::geom_pointrangeh(aes(y=Pretty, x=mean, xmin=low, xmax=high, color=gender),
                               position=position_dodge2(width=0.8, padding=0.1), size=0.25) + # ggstance is required to make legend keys horizontal
    geom_rect(data=combined_variables %>% filter(stripe==1), 
              mapping=aes(ymax = as.numeric(Pretty) + 0.5,
                  ymin = as.numeric(Pretty) - 0.5),
                  #xmin = -Inf,
                  #xmax = Inf,
                  fill = "gray", xmin=-Inf, xmax=Inf, alpha = 0.1, show.legend = FALSE, colour=NA) +
    labs(title="", #title="Effects sizes of variables on Hb prediction",
         x="Regression coefficient", y=NULL, colour="Sex") +
    scale_colour_discrete(drop=FALSE, breaks = c("male", "female")) + 
    #guides(colour = guide_legend(reverse = TRUE)) + # Reverse the order of keys in the legend
    #theme_gray(base_size = base_size) + theme(legend.position = "bottom", legend.direction = "horizontal")
    theme_classic(base_size = base_size) + theme(legend.position = "bottom", legend.direction = "horizontal")
  return(list(plot=plot, cis=cis))
}

@@

  #pretty <- descript["Pretty"]
  #pretty <- combined_variables
  rrfFit.varimp <- rrfFit.varimp %>% mutate(Pretty=factor(Pretty, levels=levels(combined_variables$Pretty)))
  rrfFit.varimp <- left_join(combined_variables, rrfFit.varimp, by="Pretty")

  var_imp <- rrfFit.varimp %>%
      mutate(Pretty=fct_rev(Pretty),
             gender=factor("both", levels=c("female", "both", "male"))) 
             #stripe=factor(row_number()%%2))
  print(var_imp) 
  df2 <<- var_imp
  varimp.plot <- var_imp %>% 
      # rrfFit.varimp %>%
      # mutate(Pretty=factor(Pretty, levels=rev(Pretty)), 
      #        stripe=factor(row_number()%%2)) %>%
    ggplot() + 
    ggstance::geom_pointrangeh(aes(y=Pretty, x=Importance, xmin=0, xmax=Importance, colour = gender), # ggstance is required to make legend keys horizontal
                               position=position_dodge2(width=0.8, padding=0.1), size=0.25, inherit.aes = FALSE) +   
                               #colour = hue_pal()(3)[2], size=0.25) +
      geom_rect(data=combined_variables %>% filter(stripe==1),
              mapping=aes(ymax = as.numeric(Pretty) + 0.5,
                    ymin = as.numeric(Pretty) - 0.5),
                    #xmin = -Inf,
                    #xmax = Inf,
                    #fill = stripe), 
              xmin=-Inf, xmax=Inf, alpha = 0.1, colour=NA, fill="gray", show.legend = FALSE, inherit.aes = FALSE) +
    labs(y=NULL, colour="Sex") + 
    guides(y="none") +
    scale_colour_discrete(drop=FALSE, breaks=c("both")) +
    #theme_gray(base_size = base_size)
    theme_classic(base_size = base_size) + theme(legend.position = "bottom", legend.direction = "horizontal")

@@

create_forest_importance_plot <- function(male_posterior, female_posterior, variables,
                                           filename=NULL,
                                           width = 180,  # width of the combined figure in millimetres
                                          base_size = 11
) {
  rrfFit_roc <- load_single("~/FRCBS/interval_prediction/results/rrfFit_roc.rdata")
  combined_variables <- tibble(Pretty=c(variables, "Life time donations", "Sex", "First Hb")) %>%
    mutate(stripe=factor(row_number()%%2),
           Pretty=factor(Pretty, levels=Pretty))
  combined_variables <<- combined_variables
  forest <- create_double_forest_plot(male_posterior, female_posterior, variables, combined_variables, base_size = base_size)$plot
  importance <- create_variable_importance_plot2(rrfFit_roc, descript, combined_variables, base_size = base_size)

@@

```{r}
finngenn_male_raw <- get_raw_result_list("finngen-male-dlmm")
finngenn_female_raw <- get_raw_result_list("finngen-female-dlmm")
male_posterior <- finngenn_male_raw$samples
female_posterior <- finngenn_female_raw$samples
variables <- finngenn_male_raw$pretty_variable_names
# Give a more descriptive name to this variable
variables <- replace(variables, variables=="Previous Hb deferral", "Previous donation deferred") #"Previous donation deferred", #"Previous Hb deferral"
```

@@

  #key_size <- as.numeric(grid::convertX(theme_get()$legend.key.size, "mm"))
  key_size <- 1.8
  #key_size <- unit(1.2, "lines")
  g <- df %>% ggplot(aes(y=Id, x=value, xmin=low, xmax=high)) + 
    ggstance::geom_pointrangeh(aes(colour=sex, shape=model)) + 
    #geom_pointrange(aes(colour=sex, fill=sex, shape=model), key_glyph = draw_key_rect) + 
    labs(x="Value", y="Data and method", colour="Sex", shape="Model") + 
    #scale_colour_manual(breaks = c("male", "female", "both"), values=c("yellow", "brown", "black"), aesthetics = c("color", "fill", "shape")) +
    scale_colour_discrete(breaks = c("male", "female", "both")) +
    scale_shape_discrete(breaks=c("lmm", "dlmm", "dt", "rf")) +
    #scale_colour_manual(values = c("female" = 1, "male" = 2, "both" = 3)) +
    guides(colour = guide_legend(override.aes = list(shape = 15, linetype=0, size=key_size))) +
    geom_blank(data=dummy) +
    facet_wrap("type", scales="free_x")
  #ggsave(filename="~/FRCBS/results-for-eba-paper/forest_plot.pdf", plot=g, dpi=600, units="mm", width=180)
  return(g)
}

@@

```{r calibration plots}
calibration_plot <- function(df, use_my_method=FALSE) {
  if (use_my_method) {
    df$cat <- cut(df$Deferred, seq(0.0, 1.0, length.out = 11))
    df2 <- df %>% group_by(cat, .drop=FALSE) %>% summarise(true_bin_prob = mean(obs=="Deferred"))
    df2 %>% ggplot(aes(cat, true_bin_prob)) + geom_point() + ylim(0, 1)
  } else {
    cal <- calibration(obs ~ Deferred, data=df, class="Deferred", cuts=11)
    ggplot(cal) + ylim(0, 100)
  }
}

calibration_plots <- function(ids) {
  data_frames <- map(ids, get_data_frame)
  names(data_frames) <- ids
  df <- bind_rows(data_frames, .id="Id")
  df <- df %>% 
    group_by(Id) %>%
    mutate(cat=cut(Deferred, seq(0.0, 1.0, length.out = 11))) %>%
    group_by(Id, cat, .drop=FALSE) %>% summarise(true_bin_prob = mean(obs=="Deferred"), count=n()) %>%
    ungroup() %>%
    mutate(Id = factor(Id, levels=ids))

  g <- df %>% ggplot(aes(cat, true_bin_prob, colour=count)) + 
    geom_point() + 
    geom_text(aes(label=count), colour="black", nudge_y=0.05) +
    ylim(0, 1.05) +
    labs(x="Predicted probability of deferral", y="True probability of deferral in each bin") +
    scale_colour_gradient(name = "Count", trans = "log10") +
    geom_abline(aes(intercept=-0.1, slope=0.1)) +
    facet_wrap("Id") +
    theme(axis.text.x = element_text(angle = 45, vjust = 0.5))

  #return(df)
  return(g)
}
```
